### 2025-01-21 pre_print/pre_print.qmd
First draft preprint

---
articletitle: |
  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members.
# shorttitle: "Doe et al. (2022)" # usually auto-generated
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{caption}
author:
  - name: James Steele
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james@steele-research.com
  - name: James Fisher
    affiliations:
      - ref: 3
    orcid: 0000-0002-6013-8402
    email: james.fisher.phd@outlook.com
  - name: Tim Dettmann
    affiliations:
      - ref: 1
    email: tim.dettmann@gmail.com
affiliations:
      - id: 1
        name: Kieser Australia
      - id: 2
        name: Steele Research Limited
      - id: 3
        name: Department of Sport and Health, Solent University
abstract: |
  Place your abstract here. Use Markdown to style the abstract, e.g. write `*word*` to print a *word* in italics. You can use the Latex command `\newline\` to create a linebreak in the abstract. \newline\ *Explanation*: The second backslash ensures that styling can be applied directly after the linebreak. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template. This document is only a demo explaining how to use the template.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [template, demo, exercise science] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: bibliography.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

@@

It has been speculated [@fisherRoleSupervisionResistance2022a] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort. Indeed, a recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022]. Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024]). Thus, whether self-selecting a load for a typically prescribed repetition range or self-selecting a load and self-selecting the number of repetitions to perform, most will perform sets with an estimated proximity of \~10-20 repetitions shy of momentary failure.

@@

We were afforded an opportunity to examine the role of supervision on effort during RT in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership[^1] or supervised by an exercise scientist on an "Assisted" membership. As such, using Bayesian methods, we examined samples of historical data from both types of client and in a sample of current Core members we investigated experimentally the impact of supervision from a qualified exercise scientist.

@@

The study was conducted at a selection of strength clinics operated by Kieser Australia recruiting from the existing pool of Core members at these locations. All Kieser Australia members are prescribed the same protocol. This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load of 90-120 seconds using a \~ 10 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 1 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action). Core trainees are prompted to progress load by \~5% for the next session once a time-under-load of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist. Trainees are prompted to complete a session rating of perceived effort (RPE) at the end of their training session. An acute randomised cross-over design was employed to examine participants completing an RT session with and without direct supervision with one week apart (i.e., a Core vs Assisted session). Participants were instructed to complete a single set of the resistance machine exercises on their current training programme card using their current training loads to momentary failure. The time under load (TUL) performed, and perceptions of effort and discomfort were recorded and compared between both conditions. This study was not pre-registered and as described below the sample size was justified based on logistical concerns and the analysis is considered exploratory.

@@

@@

We queried the Kieser Australia database to generate random samples from historical data. We opted to generate reasonably large samples to ensure precise prior estimates, though did not use the entirety of the historical data (\~50000 members, \>10 years of data) so as to reduce computation time for Bayesian modelling. For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of prior training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise. Thus we ended up with a sample of n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`. For session RPE we performed a similar database query and randomly sampled 1000 Core and 1000 Assisted members who had reported session RPE values.

@@

All RT was performed using resistance machines (Kieser Training AG, Zurich, Switzerland). For the experimental study participants were instructed to attend two training sessions at least one-week apart where they completed either their current Core session (i.e., unsupervised) or an Assisted session (i.e., supervised). During the Core session the participant completed their training session as prescribed with the only difference for the Assisted session being that an exercise scientist supervised them providing instruction and motivation, however, they did not interfere in any physical way during the set (e.g. to spot or assist in completing a failing repetition). Each participant had an existing training programme card. In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card prior to participation in the study. Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the time-under-load achieved for the purposes of the study i.e., if they realised during the set that their prior load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL. Momentary failure was defined as per Steele et al. [-@steeleClarityReportingTerminology2017] i.e., the point at which, despite their greatest effort, participants are unable to continue concentrically contracting and moving the resistance, and this was communicated to the participants. During each session for each exercise the trainees/trainers recorded their time-under-load achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose. In addition, they recorded their perceptions of effort and discomfort in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).

@@

```         
    UPDATE LINKS!!!!!!!!
```

@@

For both the prior sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept. The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used for the experimental data model only (note, due to the session RPE reported in the prior sample data and that we sampled only one session we do not have member or machine random effects, and due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset). These fixed and random effects were modelled as predictors for both the mean and precision parameters of the ordered beta regression model. As noted, RPE in both datasets was rescaled to lie on the $(0,1)$ interval. Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows:were as follows (note that the model for the prior dataset omits the random effect for machine):

@@

preds_prior_tul <- avg_predictions(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA)

preds_tul <- avg_predictions(model_tul,
                               variables = "core_assisted",
                               re_formula = NA)

contrast_prior_tul <- avg_comparisons(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA)

contrast_tul <- avg_comparisons(model_tul,
                               variables = "core_assisted",
                               re_formula = NA)

preds_prior_tul_hu <- avg_predictions(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu")

preds_tul_hu <- avg_predictions(model_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu")

contrast_prior_tul_hu <- avg_comparisons(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu")

contrast_tul_hu <- avg_comparisons(model_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu")

@@

preds_prior_rpe <- avg_predictions(model_prior_sample_rpe,
                               variables = "core_assisted",
                               re_formula = NA)

preds_rpe <- avg_predictions(model_rpe,
                               variables = "core_assisted",
                               re_formula = NA)

contrast_prior_rpe <- avg_comparisons(model_prior_sample_rpe,
                               variables = "core_assisted",
                               re_formula = NA)

contrast_rpe <- avg_comparisons(model_rpe,
                               variables = "core_assisted",
                               re_formula = NA)

preds_discomfort <- avg_predictions(model_discomfort,
                               variables = "core_assisted",
                               re_formula = NA)

contrast_discomfort <- avg_comparisons(model_discomfort,
                               variables = "core_assisted",
                               re_formula = NA)

```
In our prior sample model there was a small difference between Core and Assisted sessions in RPE. The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_prior_rpe$estimate)` [95%QI: `r scales::percent(contrast_prior_rpe$conf.low)`, `r scales::percent(contrast_prior_rpe$conf.high)`].

Our experimental sample model showed a similar magnitude of difference between Core and Assisted sessions in RPE, though the posterior estimates for RPEs were considerably greater compared with the prior sample estimates. The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_rpe$estimate)` [95%QI: `r scales::percent(contrast_rpe$conf.low)`, `r scales::percent(contrast_rpe$conf.high)`].

The rating of perceived discomfort results showed a similar pattern to the RPE albeit with a slightly larger magnitude of difference between conditions. The estimated rating of perceived discomfort as a percentage for Core sessions was `r scales::percent(preds_discomfort$estimate[2])` [95%QI: `r scales::percent(preds_discomfort$conf.low[2])`, `r scales::percent(preds_discomfort$conf.high[2])`], for Assisted sessions `r scales::percent(preds_discomfort$estimate[1])` [95%QI: `r scales::percent(preds_discomfort$conf.low[1])`, `r scales::percent(preds_discomfort$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_discomfort$estimate)` [95%QI: `r scales::percent(contrast_discomfort$conf.low)`, `r scales::percent(contrast_discomfort$conf.high)`].

@@

# Discussion

ADD DISCUSSION

>>>>

---
articletitle: |
  The effects of supervision upon effort during resistance training: A Bayesian analysis of prior data and an experimental study of private strength clinic members
# shorttitle: "Doe et al. (2022)" # usually auto-generated
format: 
  sportrxiv-pdf:
    include-in-header:
      text: |
       \usepackage[font=scriptsize]{caption}
       \usepackage{caption}
author:
  - name: James Steele
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
    orcid: 0000-0002-8003-0757
    corresponding: true
    email: james@steele-research.com
  - name: James Fisher
    affiliations:
      - ref: 3
    orcid: 0000-0002-6013-8402
    email: james.fisher.phd@outlook.com
  - name: Tim Dettmann
    affiliations:
      - ref: 1
    email: tim.dettmann@gmail.com
affiliations:
      - id: 1
        name: Kieser Australia, Melbourne, Australia
      - id: 2
        name: Steele Research Limited, Eastleigh, UK
      - id: 3
        name: Department of Sport and Health, Solent University, Southampton, UK
      - id: 4
        name: The Exercise Coach, Tennessee, USA
abstract: |
  Recent work has suggested that supervision during resistance training might enhance outcomes such as strength adaptation due to role that it plays in optimising trainee effort. We were afforded an opportunity to examine the role of supervision on effort during resistance training in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership or supervised by a qualified exercise scientist on an "Assisted" membership. Both retrospectively through examination of samples of member training records, and through a prospective experimental study, we examined the effects of supervision upon exercise performance as time under load (TUL), rating of perceived effort (RPE), and rating of perceived discomfort. We adopted Bayesian methods for modelling and from the prior sample of retrospective data generated empirically informed posterior distributions to use as prior distributions in our modelling of the experimental study. The prior sample consisted of training sessions from 1000 each of Core and Assisted members, whilst the experimental study consisted of Core members recruited to in randomised order complete an unsupervised and a supervised session using their current training loads where they were instructed to perform each exercise to momentary failure. Our findings suggest that, *in situ* under real world settings (i.e., our prior sample of members) there was little difference in exercise performance, yet in our experimental study there was a clear benefit to TUL performance when under supervision (Core = 125.12 [95%QI: 113.70, 131.90] seconds; Assisted = 147.35 [95%QI: 134.29, 154.81] seconds; Core minus Assisted contrast = -22.10 [95%QI: -26.60, -17.61] seconds). There was a small difference in RPE in percentage points reported with and without supervision under both real world (Core = 53% [95%QI: 51%, 55%]; Assisted = 59% [95%QI: 57%, 61%]; Core minus Assisted contrast = -6% [95%QI: -8%, -4%]) and experimental conditions (Core = 81% [95%QI: 75%, 86%]; Assisted = 87% [95%QI: 83%, 91%]; Core minus Assisted contrast = -6% [95%QI: -10%, -4%]) suggesting a that under supervision trainees train with greater proximity to failure, which was also supported by greater rating of perceived discomfort under supervision during the experimental study (Core = 6.3 [95%QI: 5.1, 7.3] pts; Assisted = 7.5 [95%QI: 6.5, 8.3] pts; Core minus Assisted contrast = -1.2 [95%QI: -1.6, -0.9] pts). These results in general support prior work highlighting the importance of supervision during RT and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.
license-type: ccby # change if neccessary
# year: 2025 # defaults to current year
keywords: [template, demo, exercise science] # optional
# optional link to Supplementary Material:
suppl-link: https://osf.io/ 
reference-section-title: References
printnote: "PREPRINT - NOT PEER REVIEWED" # if post print, include "POSTPRINT" then link to the published article
bibliography: bibliography.bib  
pdf-engine: xelatex
execute: 
  echo: false
  message: false
  warning: false
---

@@

It has been speculated [@fisherRoleSupervisionResistance2022a] that an explanation for this effect may be due to the role that a trainer plays in prescribing load progression for the trainee, whereas when there is no clear load progression rule unsupervised trainees may be less likely to train with appropriate loads, and ultimately appropriate effort. Indeed, a recent meta-analysis examining the relative loads selected by trainees when given the ability to self-select highlights that they tend to choose loads which, whilst initially efficacious in novice trainees, become sub-optimal quickly as training experience progresses (\~53% of one repetition maximum \[1RM\]) particularly when combined with the typical repetition ranges prescribed i.e., \~8 to 15 repetitions [@steeleAreTraineesLifting2022]. Further, when given the opportunity to self-select load and the number of repetitions to complete there is evidence that trainees likely train with relatively low effort as determined from their proximity to momentary failure i.e., \~10 repetitions at \~53% 1RM [@steeleAreTraineesLifting2022; @nuzzoMaximalNumberRepetitions2024]. Thus, whether self-selecting a load for a typically prescribed repetition range or self-selecting a load and self-selecting the number of repetitions to perform, most will perform sets with an estimated proximity of \~10-20 repetitions shy of momentary failure.

@@

We were afforded an opportunity to examine the role of supervision on effort during RT in a unique setting with several private strength clinics whose members train either unsupervised on a "Core" membership[^1] or supervised by an exercise scientist on an "Assisted" membership both retrospectively and through a prospective experimental study. As such, using Bayesian methods, we examined samples of historical data from both types of client and in a sample of current Core members in which we investigated experimentally the impact of supervision from a qualified exercise scientist.

@@

The study was conducted at a selection of strength clinics operated by Kieser Australia recruiting from the existing pool of Core members at these locations. All Kieser Australia members are prescribed the same protocol. This consists of a single set of the resistance machine exercises prescribed on their current training programme card using a load that should permit them to reach momentary failure within a time-under-load of 90-120 seconds (though an upper limit of ~180 seconds TUL is enforced to avoid machines being occupied for too long on the clinic floor preventing other members from using them) using a \~ 10 seconds repetition duration (i.e., \~4:4 seconds concentric:eccentric actions with a 1 isometric second hold whilst still under load with tension on the involved musculature at the end of each concentric and eccentric muscle action). Core trainees are prompted to progress load by \~5% for the next session once a time-under-load of \>120 seconds is achieved for a particular exercise before reaching momentary failure, and for Assisted trainees this is actioned by their supervising Exercise Scientist . Trainees are prompted to complete a session rating of perceived effort (RPE) at the end of their training session. An acute randomised cross-over design was employed to examine participants completing an RT session with and without direct supervision with one week apart (i.e., a Core vs Assisted session). Participants were instructed to complete a single set of the resistance machine exercises on their current training programme card using their current training loads to momentary failure. The time under load (TUL) performed, and perceptions of effort and discomfort were recorded and compared between both conditions. This study was not pre-registered and as described below the sample size was justified based on logistical concerns and the analysis is considered exploratory.

@@

The experimental study component of this project was approved by the Southampton Solent University Health Exercise and Sport Science Ethics committee (id: STEELEAUG2016). All participants provided informed consent to participate in the study. Data in the prior sample was used in de-identified form and with member consent for data to be used for research purposes.

@@

We queried the Kieser Australia database to generate random samples from historical data. We opted to generate reasonably large samples to ensure precise prior estimates, though did not use the entirety of the historical data (\~50000 members, \>10 years of data) so as to reduce computation time for Bayesian modelling. For TUL we limited data to Core and Assisted members training sessions that either were not led by an exercise scientist or were respectively[^3], took the first training session after at least 6 months of prior training at Kieser Australia had been completed by each member, randomly sampled 1000 Core and 1000 Assisted members and then filtered to the resistance machines used in by members in the experimental study so that we had a selection of members across varied clinic locations and completing sessions with a selection of resistance machine exercises and had TUL data for each exercise. Thus we ended up with a sample of n = `r length(unique(prior_data_tul$id))` (Core = `r length(unique(filter(prior_data_tul, core_assisted == "core")$id))`, Assisted = `r length(unique(filter(prior_data_tul, core_assisted == "assisted")$id))`). For session RPE we performed a similar database query and randomly sampled 1000 Core and 1000 Assisted members who had reported session RPE values.

@@

All RT was performed using resistance machines (Kieser Training AG, Zurich, Switzerland) including the A1 (hip extension), A2 (torso flexion), A3 (hip abduction), B1 (leg extension), B6 (leg press), B7 (seated leg curl), C3 (torso arm i.e., pulldown), C7 (seated row), D5 (arm cross i.e., pectoral fly), D6 (chest press), D7 (seated dip), F2/F2.1 (abdominal flexion), and K2 (supported supinated grip pullup). For the experimental study participants were instructed to attend two training sessions at least one-week apart where they completed either their current Core session (i.e., unsupervised) or an Assisted session (i.e., supervised). During the Core session the participant completed their training session as prescribed with the only difference for the Assisted session being that an exercise scientist supervised them providing instruction and motivation, however, they did not interfere in any physical way during the set (e.g. to spot or assist in completing a failing repetition). Each participant had an existing training programme card. In both conditions they utilised the current training loads for each machine that were recorded in the last session of their training programme card prior to participation in the study. Participants were instructed to ensure that they continue performing the exercise to momentary failure independently of the time-under-load achieved for the purposes of the study i.e., if they realised during the set that their prior load selection was evidently too low for the prescribed TUL range they should continue to momentary failure regardless off the TUL. Momentary failure was defined as per Steele et al. [-@steeleClarityReportingTerminology2017] i.e., the point at which, despite their greatest effort, participants are unable to continue concentrically contracting and moving the resistance, and this was communicated to the participants. During each session for each exercise the trainees/trainers recorded their time-under-load achieved as they would do for their usual sessions using timers situated around the clinic in view during training specifically for this purpose. In addition, they recorded their perceptions of effort and discomfort in that order immediately upon completing the exercise using previously validated scales for differentiating these perceptions [@steeleDifferentiationPerceivedEffort2017]; scripts and scales are available here: [https://osf.io/ufvy8/](https://osf.io/ufvy8/).

@@

```         
    JAMES TO UPDATE LINKS/CITATION IN PARAGRAPH BELOW!!!!!!!!
```

@@

For both the prior sample and experimental study data the following model was employed which included a fixed (i.e., population level) effect for the session type, Core or Assisted, where Assisted was coded as the intercept. The model also included random (i.e., cluster or group level) effects as intercepts for the location (i.e., the clinic the session was performed at), the member id, and the resistance machine used for the experimental data model only (note, due to the session RPE reported in the prior sample data and that we sampled only one session we do not have member or machine random effects, and due to the cross-over nature of the experimental design the random effects for location, member, and machine were modelled as nested in this dataset). These fixed and random effects were modelled as predictors for both the mean and precision parameters of the ordered beta regression model. As noted, RPE in both datasets was rescaled to lie on the $(0,1)$ interval. Formally the model(s) for $i^{th}$ RPE across condition $c$ within the $j^{th}$ location, $k^{th}$ member, $l^{th}$ machine were as follows (note that the model for the prior dataset omits the random effect for machine):

@@

preds_prior_tul <- avg_predictions(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

preds_tul <- avg_predictions(model_tul,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

contrast_prior_tul <- avg_comparisons(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

contrast_tul <- avg_comparisons(model_tul,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

preds_prior_tul_hu <- avg_predictions(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu") |>
  mutate(across(where(is.numeric), round, 2))

preds_tul_hu <- avg_predictions(model_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu") |>
  mutate(across(where(is.numeric), round, 2))

contrast_prior_tul_hu <- avg_comparisons(model_prior_sample_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu") |>
  mutate(across(where(is.numeric), round, 2))

contrast_tul_hu <- avg_comparisons(model_tul,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "hu") |>
  mutate(across(where(is.numeric), round, 2))

@@

preds_prior_rpe <- avg_predictions(model_prior_sample_rpe,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

preds_rpe <- avg_predictions(model_rpe,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

contrast_prior_rpe <- avg_comparisons(model_prior_sample_rpe,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

contrast_rpe <- avg_comparisons(model_rpe,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

preds_prior_rpe_phi <- avg_predictions(model_prior_sample_rpe,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "phi",
                               type = "link") |>
  mutate(across(where(is.numeric), round, 2))

contrast_prior_rpe_phi <- avg_comparisons(model_prior_sample_rpe,
                               variables = "core_assisted",
                               re_formula = NA,
                               dpar = "phi",
                               type = "link") |>
  mutate(across(where(is.numeric), round, 2))

preds_rpe_phi <- avg_predictions(model_rpe,
                               variables = "core_assisted",
                               re_formula = NA,
                             dpar = "phi",
                               type = "link") |>
  mutate(across(where(is.numeric), round, 2))

contrast_rpe_phi <- avg_comparisons(model_rpe,
                               variables = "core_assisted",
                               re_formula = NA,
                             dpar = "phi",
                               type = "link") |>
  mutate(across(where(is.numeric), round, 2))

preds_discomfort <- avg_predictions(model_discomfort,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

contrast_discomfort <- avg_comparisons(model_discomfort,
                               variables = "core_assisted",
                               re_formula = NA) |>
  mutate(across(where(is.numeric), round, 2))

preds_discomfort_phi <- avg_predictions(model_discomfort,
                               variables = "core_assisted",
                               re_formula = NA,
                             dpar = "phi",
                               type = "link") |>
  mutate(across(where(is.numeric), round, 2))

contrast_discomfort_phi <- avg_comparisons(model_discomfort,
                               variables = "core_assisted",
                               re_formula = NA,
                             dpar = "phi",
                               type = "link") |>
  mutate(across(where(is.numeric), round, 2))

```
In our prior sample model there was a small difference between Core and Assisted sessions in RPE. The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_prior_rpe$estimate)` [95%QI: `r scales::percent(contrast_prior_rpe$conf.low)`, `r scales::percent(contrast_prior_rpe$conf.high)`]. The $\phi$ parameter for precision of the beta distribution was higher in the Assisted condition suggesting that RPE reported was more clustered around the estimated RPE values as compared to in the Core condition. The $\phi$ parameter for Core sessions was `r preds_prior_rpe_phi$estimate[2]` [95%QI: `r preds_prior_rpe_phi$conf.low[2]`, `r preds_prior_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_prior_rpe_phi$estimate[1]` [95%QI: `r preds_prior_rpe_phi$conf.low[1]`, `r preds_prior_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_prior_rpe_phi$estimate` [95%QI: `r contrast_prior_rpe_phi$conf.low`, `r contrast_prior_rpe_phi$conf.high`].

Our experimental sample model showed a similar magnitude of difference between Core and Assisted sessions in RPE, though the posterior estimates for RPEs were considerably greater compared with the prior sample estimates. The estimated RPE as a percentage for Core sessions was `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], for Assisted sessions `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`], and the between condition contrast (Core minus Assisted) was `r scales::percent(contrast_rpe$estimate)` [95%QI: `r scales::percent(contrast_rpe$conf.low)`, `r scales::percent(contrast_rpe$conf.high)`]. Similarly to the prior data, the $\phi$ parameter for precision of the beta distribution in our experimental sample was higher in the Assisted condition suggesting that RPE reported was more clustered around the estimated RPE values as compared to in the Core condition. The $\phi$ parameter for Core sessions was `r preds_rpe_phi$estimate[2]` [95%QI: `r preds_rpe_phi$conf.low[2]`, `r preds_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_rpe_phi$estimate[1]` [95%QI: `r preds_rpe_phi$conf.low[1]`, `r preds_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_rpe_phi$estimate` [95%QI: `r contrast_rpe_phi$conf.low`, `r contrast_rpe_phi$conf.high`].

The rating of perceived discomfort results showed a similar pattern to the RPE albeit with a slightly larger magnitude of difference between conditions. The estimated rating of perceived discomfort as a percentage for Core sessions was `r preds_discomfort$estimate[2]*10` [95%QI: `r preds_discomfort$conf.low[2]*10`, `r preds_discomfort$conf.high[2]*10`], for Assisted sessions `r preds_discomfort$estimate[1]*10` [95%QI: `r preds_discomfort$conf.low[1]*10`, `r preds_discomfort$conf.high[1]*10`], and the between condition contrast (Core minus Assisted) was `r contrast_discomfort$estimate*10` [95%QI: `r contrast_discomfort$conf.low*10`, `r contrast_discomfort$conf.high*10`]. The $\phi$ parameter for precision of the beta distribution was higher in the Core condition suggesting that rating of perceived discomfort reported was more clustered around the estimated rating of perceived discomfort values as compared to in the Assisted condition, though the contrast for this was less precise than in the RPE models. The $\phi$ parameter for Core sessions was `r preds_prior_rpe_phi$estimate[2]` [95%QI: `r preds_prior_rpe_phi$conf.low[2]`, `r preds_prior_rpe_phi$conf.high[2]`], for Assisted sessions `r preds_prior_rpe_phi$estimate[1]` [95%QI: `r preds_prior_rpe_phi$conf.low[1]`, `r preds_prior_rpe_phi$conf.high[1]`], and the between condition contrast (Core minus Assisted) was `r contrast_prior_rpe_phi$estimate` [95%QI: `r contrast_prior_rpe_phi$conf.low`, `r contrast_prior_rpe_phi$conf.high`].

@@

# Discussion
Our study is a first to offer insights into the role of supervision (i.e., Core vs Assisted sessions) during RT both *in situ* in an ecologically valid real world setting in addition to during an experimental study. Under standardised training protocol prescription (i.e., single sets of RT using resistance machines to momentary failure using a fixed repetition duration using a load that should permit momentary failure in a target TUL range of 90-120 seconds) we examined exercise performance as the TUL, in addition to RPE and rating of perceived discomfort. Our key findings were that under real world settings in our prior sample there did not appear to be a difference in TUL between Core and Assisted sessions nor the probability of whether members targeted the upper bound of the prescribed TUL range (i.e., 120s), yet in our experimental study there was a clear benefit to TUL performance when under supervision and reduction in the probability of targeting the upper prescribed TUL range. There was a small difference in RPE reported with and without supervision under both real world and experimental conditions with members reporting higher RPE in Assisted training conditions, though RPE on the percentage points scale was notably higher during the experimental study compared with the prior sample for both Core and Assisted sessions. Lastly, rating of perceived discomfort was also higher in the experimental study under supervised Assisted sessions compared to Core sessions. These results in general support prior work highlighting the importance of supervision during RT [@fisherRoleSupervisionResistance2022a; @fisherSupervisionResistanceTraining2023; @carlsonMaleFemalePerceptions2024] and that under unsupervised, and perhaps more so during real world conditions, trainees likely train with suboptimal effort.

We will first discuss the findings from the prior sample of Core and Assisted members. In this *in situ* dataset we find for both Core and Assisted members there is clear evidence of targeting a specific TUL based on the prescribed TUL range of 90-120 seconds. For Core members this is perhaps unsurprising given that, in repetition count based RT prescription, when given the opportunity to self-select repetition number people tend to target ~10 repetitions regardless of the load they self-select and that also most RT prescriptions tend to emphasise that particular target [@steeleAreTraineesLifting2022]. Thus, that members are also shooting for targets under this slightly different type of RT protocol suggests that this might be characteristic of most people training under real world unsupervised conditions. It is however surprising that this is also the case for Assisted members and that TUL differs very little between Core and Assisted sessions. However, the lack of difference in TUL itself is not necessarily indicative of suboptimal training effort, and we did observe slightly higher reported RPE during Assisted training sessions compared to Core sessions. Thus it might be that during Assisted sessions trainees might be more likely to use loads more appropriate for the prescribed TUL range and thus train with closer proximity to failure as compared to under Core sessions in addition to greater clustering of RPE around higher values (indicated by the $\phi$ parameter in the ordered beta regression). This may be due to load progression protocols being more closely followed by supervising exercise scientists as compared to those training unsupervised where, unless using initially heavy loads or in previously well trained persons, most unsupervised trainees do not progress load sufficiently [@steeleAreTraineesLifting2022; @fisherRoleSupervisionResistance2022a]. In future work we plan to explore differences longitudinally in load progression between Core and Assisted members. But, despite the greater RPE reported under Assisted sessions, in percentage points RPE under both Core and Assisted sessions was considerably lower in our prior sample (Core = `r scales::percent(preds_prior_rpe$estimate[2])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[2])`, `r scales::percent(preds_prior_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_prior_rpe$estimate[1])` [95%QI: `r scales::percent(preds_prior_rpe$conf.low[1])`, `r scales::percent(preds_prior_rpe$conf.high[1])`]) compared with our experimental study (Core = `r scales::percent(preds_rpe$estimate[2])` [95%QI: `r scales::percent(preds_rpe$conf.low[2])`, `r scales::percent(preds_rpe$conf.high[2])`], Assisted = `r scales::percent(preds_rpe$estimate[1])` [95%QI: `r scales::percent(preds_rpe$conf.low[1])`, `r scales::percent(preds_rpe$conf.high[1])`]) suggesting that members in these real world conditions may not be training with sufficient effort and many perhaps "going through the motions" [@steeleHigherEffortbasedParadigm2017].
